---
title: "Investigating Global Plastic Emissions in the Ocean"
author: "Team Diamond Ocean: Purumidha Sharma, Sanjana Girish, Stephanie Cristea, Anita Zabojnikova, Andrew Gritsevskiy, and Danur Mahendra"
output: html_document
runtime: shiny
date: "Date Published: 16th November, 2022"
---

<style type="text/css">

p {
    font-size: 14px;
    line-height: 20px;
    margin: 0px 0px 12px 0px;
}

h1, h2, h3, h4, h5, h6, legend {
    font-family: Arial, sans-serif;
    font-size: 23px;
    font-weight: 20;
    color: #8097d3;
}


h1.title {
  font-size: 38px;
  font-style: bold;
  color: darkBlue;
  text-align: center;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: darkBlue;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 13px;
  font-family: "Times New Roman", Times, serif;
  color: blue;
  text-align: center;
}
</style>


```{r, echo=FALSE, warning = FALSE, message = FALSE}
# Loading Libraries -- please put them all here, so we don't have duplicates 
#install.packages("cartogram")
library(RColorBrewer)
library(maptools)
library(ggmap)
library(rgeos)
library(sf)
library(rgdal)
library(cartogram)
require(rgdal)
library(raster)
library(dplyr)
library(tidyverse)
library(lineupjs)
library(shiny)
library(shinyWidgets)
library(crosstalk)
library(magrittr)
library(readxl)
library(sf)
library(spData)
library(ggpattern)
library(readr)
library(png)
library(magick)
```

```{r setup, echo=FALSE, include=FALSE}
# To collapse warnings and messages 
knitr::opts_chunk$set(message=FALSE, warning=FALSE, 
                     error=FALSE)
```

Plastic has become a vital aspect in our everyday life. Its extensive range of applications allows for economic growth and is therefore a crucial component to many industrial and commercial purposes. However, most types of plastic contain non-biodegradable materials which require additional care for proper disposal. Wealthier nations who can afford to offload their share of the burden often do so at the expense of developing nations. In many cases, developing nations import plastic waste out of necessity in exchange for monetary benefits but do not have the capacity for a proper disposal, meaning a significant portion gets expelled to the ocean. This presents the image that developing nations, particularly in the South-Eastern Asia area, are ultimately responsible for the majority of ocean waste. In this data-driven article, we will first examine the origins of ocean plastic and how nations use international trade as a means to dispose of their plastic emission. We will present innovative new technologies on the rise, as well as some of the best waste management policies that corporations and global leaders can mimic. Ultimately, plastic emissions has a global effect, and it is in the best interest of everyone to minimize plastic waste expelled to the ocean. 


## Which Countries Contribute the Most Ocean Plastic?

<iframe src="https://public.tableau.com/views/plasticwasteemissiontotheoceanbycountry2019/Sheet1?:language=en-US&:display_count=n&:origin=viz_share_link"></iframe>
              
The visualization above provides more detail on which country contributes plastic to the ocean. By far, the top producer of plastic waste that ends up in the ocean is the Philippines, producing 356,371 tons. This is almost three times the amount of the second highest producer, India. About 10 countries produce the bulk of the world’s plastic waste in oceans, most of which are in Asia. Only two of these countries are in the top ten for plastic waste production per capita: the Philippines and Malaysia. Most of the countries that have the highest waste per capita are tiny, developing countries such as Suriname, Saint Lucia, and Grenada, and have negligible total waste. There is not much of a relationship between total emissions and per capita emissions. As a result, a natural supposition is that the countries who are in most need of a waste emission and production intervention. However, this fails to account for plastic waste that is imported by these countries. The following visualizations will attempt to investigate where these plastic emissions are actually coming from. 

## How do plastic imports and exports affect countries' ultimate responsibilities for plastic pollution?

<iframe src="https://public.tableau.com/shared/9Y8JTWJY5?:display_count=n&:origin=viz_share_link"></iframe>
 
There is a significant difference in how much plastic is imported and exported by country. As is visible from the visualization above, North American and European countries are responsible for a highly disproportionate number of plastic waste exports, the majority of which end up in countries such as Malaysia, India, and the Philippines. 

<iframe src="https://public.tableau.com/views/test_16711405742330/Dashboard1?:language=en-US&:display_count=n&:origin=viz_share_link"></iframe>

The figure above demonstrates the trend between exported plastic and plastic that ends up in the ocean. The countries that export the most plastic are also the ones that do not expel much plastic into the ocean (for instance, the United States and China). This is unsurprising because they have a finite amount of plastic, and there is no need to expel it into the ocean, if they are exporting it. However, on the other end of the spectrum, there are countries that export very little of their plastic which expel more into the ocean (for instance the Philippines, and Suriname). This suggests that plastic expelled into the ocean by low-export countries is most likely to be plastic exported by high-export countries. 

As such, it is important for wealthier countries to recognize the extent to which they contribute to global plastic waste, even though on a country-level scale, it may seem as if plastic is being properly disposed of. In fact, it is a shared global responsibility to ensure that plastic is properly disposed of, instead of being exported to countries which will dispose of it into the ocean. Several steps can be taken to reduce plastic waste and improve waste management systems which will be further discussed below. 


## Have innovations in plastic reducing technologies increased lately? Are there emerging technologies that can help clean plastic waste in oceans? 

So far, we have only mentioned problems related to global plastic emissions in ocean, however, let us now delve into possible innovative solutions to stop our world from being a plastic planet. 

```{r, echo=FALSE}
# VISUALIZATION 4: HAVE INNOVATIONS IN PLASTIC REDUCTING TECHNOLOGIES INCREASED LATELY? 

innovation_patents <- read_excel("new invention patents from 1970-2017 newxlsx.xlsx")

.lineupDefaultOptions <- list(
  filterGlobally = TRUE,
  singleSelection = FALSE,
  animated = TRUE,
  sidePanel = "collapsed",
  hierarchyIndicator = TRUE,
  labelRotation = 0,
  rowHeight = 18,
  rowPadding = 2,
  groupHeight = 40,
  groupPadding = 5,
  summaryHeader = TRUE,
  overviewMode = TRUE,
  expandLineOnHover = TRUE,
  defaultSlopeGraphMode = "item",
  ignoreUnsupportedBrowser = FALSE
)
# clean the data: 

innovation_patents_new <- innovation_patents %>% 
                              janitor::clean_names()

#new_innovation_patents_redefine <- innovation_patents_new %>% group_by(waste_management_type) %>% 
  #summarise(n = n(), sum_num = sum(worldwide_number_of_patented_innovations_in_plastic_reducing_technologies)) %>%
 # mutate(average_patented_innovations = round(sum_num/n))

#new_data <- left_join(innovation_patents_redefine,new_innovation_patents_redefine )


innovation_patents_cleaned <- innovation_patents_new %>% 
  `row.names<-`(., NULL) %>% 
  pivot_wider(names_from = waste_management_type,
              values_from = worldwide_number_of_patented_innovations_in_plastic_reducing_technologies) %>% column_to_rownames(var = "year") 

# Define UI for application that draws a histogram
ui <- fluidPage(
  titlePanel("Worldwide Patented Innovations in Plastic Reducing Technologies"),
  fluidRow(column(12, align = 'top',
                  wellPanel(style = "background-color: #D0ECFA; border-color: #2c3e50; border-size: 10px; height: 193px; width: 725px",
                            p(tags$b("Average Patented Innovations in Plastic Reducing Technologies (1990-2017)", style = "font-size: 80%")),
                            p(em(tags$b("Prevention and Recycling: 513", style = "font-size: 65%"))),
                            p(em(tags$b("Biobased feedstock: 318", style = "font-size: 65%"))),
                            p(em(tags$b("Waste Prevention: 287", style = "font-size: 65%"))),
                            p(em(tags$b("Waste Recycling: 226", style = "font-size: 65%"))),
                             p(em(tags$b("Conversion/disposal of waste and leakage removal: 197", style = "font-size: 65%")))
                           ))),
  fluidRow(column(10, align = "left", lineupOutput("lineup1"),
           tags$head(
             HTML(
               "
          <script>
          var socket_timeout_interval
          var n = 0
          $(document).on('shiny:connected', function(event) {
          socket_timeout_interval = setInterval(function(){
          Shiny.onInputChange('count', n++)
          }, 15000)
          });
          $(document).on('shiny:disconnected', function(event) {
          clearInterval(socket_timeout_interval)
          });
          </script>
          "
             )
           ), textOutput("keepAlive"),
    tags$head(tags$style(type="text/css", "label{ display: table-cell; text-align: center; vertical-align: middle; } .form-group { display: table-row;}")
    ), tags$h2("New Innovation Technologies: "),
    setBackgroundImage(
      src = "https://i.pinimg.com/originals/6b/8d/09/6b8d09037774b132e1f1c9ad50b70c92.jpg"),
    p(strong(code("Prevention and Recycling: Collecting leftover plastic to be used in road construction/bike pathway technologies"))), 
    p(strong(code("Waste Recycling: Using AI generated fungal enzyme to recycle/reduce the hardest of plastics in oceans"))), 
    p(strong(code("Waste Prevention: Having Government run incentive programs so that businesses can leverage creativity to prevent plastic use, and fine heavily for rule breakers"))),
    p(strong(code("Biobased Feedstock: BP Infinia technology to turn difficult plastic waste into recycled feedstock, Using Technology to make Biobased polymers which is from agricultural byproducts or feedstocks subtitute"))),
    p(strong(code("Conversion or Disposal of Waste and Leakage Removal: Cat-HTR Technology uses heat/pressure to convert waste plastic back into the chemical and oils it was made from")))
  )))

# Define server logic required to draw a histogram
server <- function(input, output) {
  shared_patent_innovation <- SharedData$new(innovation_patents_cleaned)
  
  output$keepAlive <- renderText({
    req(input$count)
    paste("", input$count)
  })
  
  
  output$lineup1 <- renderLineup({
    lineup(shared_patent_innovation, width = "60%", options = c(.lineupDefaultOptions), 
           ranking = lineupRanking(columns = c("*")))

  })  }
 


# Run the application
shinyApp(ui = ui, server = server)

```

As we can see here, the visualization above emphasizes how innovations in technologies that reduce plastic waste are increasing in modern times compared to the 1990s. It also helps in decoding areas, such as, plastic conversion/leakage removal, that need improvements in innovations (has least amount of innovations: 197) so that we can create a circular economy (eliminate the concept of waste). If we look at the best invested innovation sectors in waste management types, combined Prevention and Recycling wins it! In fact, this is very important because if we solve the root cause of plastic waste pollution, none of it will dwell into the oceans. 

Next, we need to understand what technologies are being invented in waste management types, such as waste prevention, to combat plastic waste in ocean? As mentioned at the bottom of the visualization, there are new technologies such as CAT-HTR which converts all types of plastic back into the chemicals and oil it was composed of. We know that conversion of plastic/leakage removal has lowest innovations, however, this one technology can revolutionize this sector in plastic reduction. To create a compelling plan of action in plastic waste management, we need to look into the countries which have best waste management strategies/policies so that other countries could mimic that. 


## Where are some of the best waste management techniques placed? How can we reciprocate the best waste management strategies from countries that handle it the best? 
```{r, echo=FALSE}
# VISUALIZATION 5: WHERE ARE SOME OF THE BEST WASTE MANAGEMENT TECHNIQUES PLACED? SANJANA

graph6_data <- read_csv("Belgium recovers recycles everything municipal waste.csv")

# Define UI for application that draws a histogram
ui <- fluidPage(

    # Application title
    titlePanel("Waste Treatment Strategies Across Countries"),
    sidebarLayout(position = "right",
        sidebarPanel(radioButtons(inputId = "wts", 
                                  "Waste Treatment Strategy",
                                  choiceNames = c("Recycling", 
                                                  "Landfill",
                                                  "Composting", 
                                                  "Incineration With Energy Recovery"),
                                  choiceValues=c("Recycling", 
                                            "Landfill",
                                            "Composting", 
                                            "Incineration_with_energy_recovery")),
                     selectInput(inputId= "choiceSelectize", 
                                 "Select Countries:", 
                                 choices = graph6_data$Country, 
                                 multiple = TRUE),
                     selectInput(inputId="AscOrDesc", "Order Bars by", 
                                 choices = c("Ascending", "Descending")),
                     selectInput(inputId="SelCountry", "Find out more about:", 
                                 choices = graph6_data$Country),
                     strong("Country:"),
                     em(textOutput("CountryText", inline=T)),
                     br(),
                     strong("Waste Management Policy:"),
                     a(textOutput("WMPURL")),
                     br(),
                     strong("Recycling Program:"),
                     a(textOutput("RPURL"))),
        # Show a plot of the generated distribution
        mainPanel( plotOutput(outputId = "graph6",  width = "100%")
        )
    )
)

# Define server logic required to draw a histogram
server <- function(input, output) {
    
    # subset countries and select necessary columns
    output$graph6 <- renderPlot({
        if (is.null(input$choiceSelectize)) {
            filter_country <- graph6_data %>%
            dplyr::select(Country, input$wts)
        } else {filter_country <- graph6_data %>% 
                filter(Country %in% input$choiceSelectize) %>%
                dplyr::select(Country, input$wts)}

        # change attributes specific to each waste management category 
        if (input$wts == "Incineration_with_energy_recovery"){
            y_label <- "Incineration With Energy Recovery Rate (in %)"
            bar_fill_col <- "#ffcccb"
            y_limit <- c(0, 80)
            y_breaks <- seq(0, 80, by = 5)
            img <- "Incineration.png"
        } else if (input$wts == "Recycling"){ 
            y_label <- paste(input$wts, "Rate (in %)")
            bar_fill_col <- "#d2f8d2"
            y_limit <- c(0, 65)
            img <- "Recycle.png"
            y_breaks <- seq(0, 65, by = 5)
        } else if (input$wts == "Landfill"){
            y_label <- paste(input$wts, "Rate (in %)")
            bar_fill_col <- "#d8cbc4"
            y_limit <- c(0, 110)
            y_breaks <- seq(0, 110, by = 5)
            img <- "Landfill.png"
        } else{ # composting
            y_label <- paste(input$wts, "Rate (in %)")
            bar_fill_col <- "#ADD8E6"
            y_limit <- c(0, 37)
            y_breaks <- seq(0, 37, by = 5)
            img <- "Composting-3.png"
        }
        
        # ascending and descending
        if (input$AscOrDesc == "Ascending"){
            filter_country$Country <- reorder(filter_country$Country, 
                                              filter_country[[input$wts]])
        } else { # Descending
            filter_country$Country <- reorder(filter_country$Country, 
                                              -filter_country[[input$wts]])
        }

        # plot the graph
        ggplot(filter_country, aes(x=Country)) +
            geom_bar_pattern(stat = "identity", aes_string(y=input$wts),
                             fill = "white", colour=bar_fill_col,
                             width=0.8, position = position_dodge(width=1), 
                             pattern= 'image', pattern_type='tile',
                             pattern_filename = img, pattern_scale=0.1) + 
            coord_flip() +
            theme(axis.text = element_text(size=11), 
                  axis.title = element_text(size=12, face="bold")) +
            labs(x = "Country", y = y_label) + 
            geom_label(aes_string(y = input$wts, label = input$wts), hjust=-0.35,
                       fill=bar_fill_col) + 
            scale_y_continuous(limits=y_limit, breaks = y_breaks) + 
            scale_pattern_discrete(guide = guide_legend(nrow = 1))
        
        # set the height and width of this ggplot
    }, height = 800, width = 610 )
    
    output$CountryText <- renderText({
        paste0(input$SelCountry)
    })
    
    output$WMPURL <- renderText({
        data_row <- graph6_data %>% filter(Country == input$SelCountry)
        (paste0(data_row[9]))
    })
    
    output$RPURL <- renderText({
        data_row <- graph6_data %>% filter(Country == input$SelCountry)
        (paste0(data_row[10]))
    })
    
    }


# Run the application 
shinyApp(ui = ui, server = server)
```

From this visualization, we can see that leading countries in recycling having clear policy objectives, bigger government funding to invest in emerging technologies like CAT-HTR and focus on producer responsibility. However, there is no country that is excelling in all waste management strategies. Some countries focus more on incinerating waste, and funding is focused more towards waste minimization rather than recovery. Other countries allocate a higher budget towards waste recovery, implementing interesting techniques like smart waste bins, pneumatic waste pipes E-waste kiosks. In terms of the best policies, we can see above that Germany has better waste management strategies compared to United States. Through research, it's found that Germany has a policy called "Green Dot" which revolves around the crux that manufacturers and retailers will need to pay more if they are using plastics anywhere in their products. This leads to less plastic waste, and therefore creating less garbage to be recycled. The net result: "a drastic decline of about one million tons less garbage than normal every year" (*All about recycling in Germany* 2022). It is surprising to see that Korea has the best waste management strategies despite highly powerful countries which breed on innovation and technology such as, United States. Korea has a "pay-to-throw" policy which is if you have bought plastic, and need to dispose it, you will need to pay thrice the amount you bought it in for. This discourages use of plastic in the first place. These are some of the most effective waste management strategies from countries that can handle it the best. This policies should be mimicked in countries that are not able to manage plastic waste properly, such as, America, Canada, and Turkey.

## Limitations and Future Considerations  

There are some limitations that are important to keep in mind when viewing the visualizations. For the visualization that examines which countries emit plastic into the ocean, the scales for the left and right x-axes are different, while a reader may make an initial assumption that they are the same and the values are comparable. The second visualization normalizes the imports and exports so that they all fit on the graph, but this means the y-axis has no relevant meaning and does not provide information on the amounts that are produced and exported. Further, there are so many countries that neither produce nor export much plastic that the lines at the bottom of the graph are difficult to distinguish. Given a chance, we would want to scale it, and perhaps try to subset our data to make it distinguishable. Furthermore, the bubble chart to depict exports and imports from countries might be a little hard to distinguish because of small bubbles not being visible. This problem of visibility could be solved by choosing a more sophisticated map, such as, honeycomb plot (distortion map) which helps to scale by imports and exports. This would be done with the use of more data since data availability was limited. 


## Key Takeaways

Based on our findings, we determined that although the majority of global plastic emission comes from developing nations particularly in southeast Asia, it does not depict an accurate picture as to who are ultimately responsible for the ever-increasing global plastic emission. 

We found that North American and European countries tend to export the majority of their emissions to Asian countries as opposed to self-disposal. Similarly, emissions from countries such as Malaysia and the Philippines tend to originate from foreign nations who export their waste to willing nations in exchange for economic gain. For instance, developing nations in the SE Asia region release far more plastic emission than any other region on Earth. However, they neither have the economic means nor population to justify such emission. We then discovered that the majority of their plastic emission was imported from other nations. We found that wealthier nations out-produced developing nations in plastic emission but chose to export their waste. By nature of being underdeveloped, developing nations may not have the means or are inefficient in plastic waste disposal but nevertheless accept imports for economic gain. 

Fortunately with modern technology, there are multiple solutions available for nations who generate high emission and/or import waste. For example, wealthier nations with large government funding may implement the use of modern technologies such as CAT-HTR to reduce their mismanaged waste. Furthermore, they can try to mimic exemplary waste management systems from leading nations for a more efficient measure to reduce their emission. However, we acknowledge that some developing nations may face financial limitations and thus, are unable to fully imitate model nations. These come from a combination of upcoming technologies, as well as strong plastic-free policies as mentioned above. 


\newpage 

## References 

All about recycling in Germany. How To Germany. (2022). Retrieved December 16, 2022, from https://www.howtogermany.com/pages/recycling.html#:~:text=Germany%20produces%2030%20million%20tons%20of%20garbage%20annually.,more%20packaging%20there%20is%2C%20the%20higher%20the%20fee.

BigRentz, BigRentz. “Bigrentz.” BigRentz, 15 Oct. 2021, https://www.bigrentz.com/blog/smart-waste-management. 

City of Toronto. “Solid Waste Reports &amp; Diversion Rates.” City of Toronto, 11 Apr. 2022, https://www.toronto.ca/services-payments/recycling-organics-garbage/solid-waste-reports/. 

Epa, Nsw. “The Waste Hierarchy.” NSW Environment Protection Authority, 20 Oct. 2022, https://www.epa.nsw.gov.au/your-environment/recycling-and-reuse/warr-strategy/the-waste-hierarchy. 

"Global Plastics Outlook: Worldwide patented inventions in environmentally relevant plastics technologies, 1990-2017." OECDiLibrary, 22 Feb. 2022, www.oecd-ilibrary.org/environment/innovation-in-plastics-prevention-and-recycling-have-grown-the-most_3e4e50b4-en.

Hazlegreaves, Steph. “Which Countries Are the Most Efficient at Minimising Waste?” Open Access Government, 19 Oct. 2020, https://www.openaccessgovernment.org/which-countries-are-the-most-efficient-at-minimising-waste/96037/. 

Lowe, Andrea. “Increasing the Timeout for Shiny Server – Domino Support.” Domino, 2020, https://tickets.dominodatalab.com/hc/en-us/articles/360058494751-Increasing-the-timeout-for-Shiny-Server.

“Plastic or Rubber Exports by Country.” Plastic or Rubber Exports by Country US$000 2020 | WITS Data, https://wits.worldbank.org/CountryProfile/en/Country/WLD/Year/LTST/TradeFlow/Export/Partner/by-country/Product/39-40_PlastiRub#. 

Ritchie, Hannah, and Max Roser. “Plastic Pollution.” Our World in Data, 1 Sept. 2018, https://ourworldindata.org/plastic-pollution. 





